```{r setup, include=FALSE}
library(Seurat)
library(dplyr)
library(sqldf)
library(ggplot2)
library(patchwork)
library(RColorBrewer)
library(ggpubr)
library(spatstat)
library(gplots)
library(corrplot)
library(stringr)

library(VennDiagram)
library(grid)#for printing PDFs off finder makrer
DiscretePalette(40, palette = "polychrome")
library(reshape2)
library(ggplot2)
options(ggrepel.max.overlaps = Inf)
#PLEASE ADD YOUR WORKING PATH HERE
study_path<-"your_path_here/"
#CREATE A FOLDER WITHIN 'study_path' as 'R'. ALL ANALYSIS WILL BE STORED HERE
save_path<-paste0(study_path,"R/")
```
READ RDS
```{r readRDS}
cosmx<-readRDS(paste0(study_path, "SL2_seurat_object.Rds"))
```
RENAME BASED ON MARKER GENES PER IDENTITY ("nb_clus") PROVIDED BY NANOSTRING 
```{r find markers and reset idents}
Idents(cosmx) <- "nb_clus"
cosmx_nb_clus_markers <- FindAllMarkers(cosmx, only.pos = T)
# Rename identity classes based on marker genes. These were previously not identified by Nanostring
cosmx <- RenameIdents(object = cosmx, c( "a" = "Mac_rich", "b" = "Igg_rich", "c"= "fib_rich", "d" = "ker_1", "e" = "ker_2", "f" ="ker_3", "neutrophil" = "neutrophil_rich", "T CD4 memory" = "T_CD4_mem",	"T CD4 naive" = "T_CD4_naive",	"T CD8 memory" = "T_CD8_mem",	"T CD8 naive" = "T_CD8_naive", "monocyte" = "inf_mac2"))
levels(cosmx)
#Hard rename nb_clus
cosmx$nb_clus<-Idents(cosmx)
```
SUB_CLUSTER INDIVIDUAL IDENTITIES FROM ABOVE TO FIND GRANULAR PHENOTYPES/CELL IDENTITIES - "Mac_rich"
```{r}
#subset
mac_seurat1 <- subset(x = cosmx, idents = c("Mac_rich"))
#run pca
mac_seurat1 <- RunPCA(mac_seurat1, assay = "SCT", verbose = FALSE)
#check pca characteristics
ElbowPlot(mac_seurat1)
#re-cluster
mac_seurat1 <- FindNeighbors(mac_seurat1, reduction = "pca", dims = 1:50)
mac_seurat1 <- FindClusters(mac_seurat1, verbose = FALSE, resolution = 0.8)
#run umap for visualisation
mac_seurat1 <- RunUMAP(mac_seurat1, reduction = "pca", dims = 1:50)
#this pdf is not used in any figures but shows how granularity within "Mac_rich" identity
pdf(paste0(save_path, "mac_seurat1_DimPlot_PC1to50_0.8res.pdf"))
  DimPlot(mac_seurat1, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
#run to identify marker genes within "Mac_rich" sub-clusters
mac_seurat1.markers <- FindAllMarkers(mac_seurat1, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(mac_seurat1.markers, paste0(save_path,"mac_seurat1_de_markers_pc50res0.8.csv"), row.names = TRUE)
#assign new names based on gene expression as identified above
DefaultAssay(mac_seurat1)<-"SCT"
mac_seurat1<-SetIdent(mac_seurat1, value = mac_seurat1$seurat_clusters)
new.cluster.ids <- c("mix","CHI3L1_mac","moDC1","inf_mac2","inf_mac1","Treg","inf_mac1","DC1","MT2A_mac","moDC1","inf_mac2","inf_mac1","MT2A_mac","DC3","Neutrophil","DC3","DC1","inf_mac2","MT2A_mac","inf_mac1","DC1","Neutrophil","mix","inf_mac2","inf_mac2","platelets","mix","inf_mac2","MT2A_mac","Neutrophil","MT2A_mac")
#create new meta data column morpho to retain new names
names(new.cluster.ids) <- levels(mac_seurat1)
mac_seurat1 <- RenameIdents(mac_seurat1, new.cluster.ids)
mac_seurat1$morpho <- Idents(mac_seurat1)
levels(Idents(mac_seurat1))
#this pdf is not used in any figures but shows how granularity within "Mac_rich" identity
pdf(paste0(save_path, "mac_seurat1_DimPlot_PC1to50_0.8res_labelled.pdf"))
  DimPlot(mac_seurat1, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()

```
SUB_CLUSTER INDIVIDUAL IDENTITIES FROM ABOVE TO FIND GRANULAR PHENOTYPES/CELL IDENTITIES - "mDC" 
```{r}
#subset
mDC_seurat <- subset(x = cosmx, idents = c("mDC"))
#run pca
mDC_seurat <- RunPCA(mDC_seurat, assay = "SCT", verbose = FALSE)
#check pca characteristics
ElbowPlot(mDC_seurat)
#re-cluster
mDC_seurat <- FindNeighbors(mDC_seurat, reduction = "pca", dims = 1:50)
mDC_seurat <- FindClusters(mDC_seurat, verbose = FALSE, resolution = 0.8)
#run umap for visualisation
mDC_seurat <- RunUMAP(mDC_seurat, reduction = "pca", dims = 1:50)
#this pdf is not used in any figures but shows how granularity within "mDC" identity
pdf(paste0(save_path, "mDC_seurat_DimPlot_PC1to20_0.8res.pdf"))
  DimPlot(mDC_seurat, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
#run to identify marker genes within "mDC" sub-clusters
mDC_seurat.markers <- FindAllMarkers(mDC_seurat, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(mDC_seurat.markers, paste0(save_path,"mDC_seurat_de_markers_pc20res0.8.csv"), row.names = TRUE)
#assign new names based on gene expression as identified above
DefaultAssay(mDC_seurat)<-"SCT"
mDC_seurat<-SetIdent(mDC_seurat, value = mDC_seurat$seurat_clusters)
new.cluster.ids <- c("mix","DC1","DC2","DC2","DC3","DC2","DC1","res_mac","mix","mix")
#create new meta data column morpho to retain new names
names(new.cluster.ids) <- levels(mDC_seurat)
mDC_seurat <- RenameIdents(mDC_seurat, new.cluster.ids)
mDC_seurat$morpho <- Idents(mDC_seurat)
levels(Idents(mDC_seurat))
#this pdf is not used in any figures but shows how granularity within "mDC" identity
pdf(paste0(save_path, "mDC_seurat_DimPlot_PC1to50_0.8res_labelled.pdf"))
  DimPlot(mDC_seurat, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
```
SUB_CLUSTER INDIVIDUAL IDENTITIES FROM ABOVE TO FIND GRANULAR PHENOTYPES/CELL IDENTITIES - "Igg_rich" 
```{r}
#subset
igg_seurat <- subset(x = cosmx, idents = c("Igg_rich"))
#run pca
igg_seurat <- RunPCA(igg_seurat, assay = "SCT", verbose = FALSE)
#check pca characteristics  
ElbowPlot(igg_seurat)
#re-cluster
igg_seurat <- FindNeighbors(igg_seurat, reduction = "pca", dims = 1:50)
igg_seurat <- FindClusters(igg_seurat, verbose = FALSE, resolution = 1.2)
#run umap for visualisation
igg_seurat <- RunUMAP(igg_seurat, reduction = "pca", dims = 1:50)
#this pdf is not used in any figures but shows how granularity within "Igg_rich" identity
pdf(paste0(save_path, "igg_seurat_DimPlot_PC1to50_0.8res.pdf"))
  DimPlot(igg_seurat, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
#run to identify marker genes within "Igg_rich" sub-clusters
igg_seurat.markers <- FindAllMarkers(igg_seurat, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(igg_seurat.markers, paste0(save_path,"igg_seurat_de_markers_pc50res0.8.csv"), row.names = TRUE)
#assign new names based on gene expression as identified above
DefaultAssay(igg_seurat)<-"SCT"
igg_seurat<-SetIdent(igg_seurat, value = igg_seurat$seurat_clusters)
new.cluster.ids <- c("mix","plasmablast","plasmablast","fibroblast3_igg","NK_igg","plasmablast","mix","plasmablast","plasmablast","mix","mix","fibroblast1_igg","mix","plasmablast","neutrophil_igg","mix","mix","mix","mix","plasmablast","plasmablast","mix","plasmablast","plasmablast","LC2_igg","NK_igg","LC2_igg","mix","plasmablast","mix","plasmablast","plasmablast","plasmablast_cycl","mast_igg")
#create new meta data column morpho to retain new names
names(new.cluster.ids) <- levels(igg_seurat)
igg_seurat <- RenameIdents(igg_seurat, new.cluster.ids)
igg_seurat$morpho <- Idents(igg_seurat)
levels(Idents(igg_seurat))
#this pdf is not used in any figures but shows how granularity within "Igg_rich" identity
pdf(paste0(save_path, "igg_seurat_labelled_DimPlot_PC1to50_0.8res_edit.pdf"))
  DimPlot(igg_seurat, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
```
SUB_CLUSTER INDIVIDUAL IDENTITIES FROM ABOVE TO FIND GRANULAR PHENOTYPES/CELL IDENTITIES - "fib_rich" 
```{r subclustering unsupervised clusters}
#subset
fib_seurat1 <- subset(x = cosmx, idents = c("fib_rich"))
#run pca
fib_seurat1 <- RunPCA(fib_seurat1, assay = "SCT", verbose = FALSE)
#check pca characteristics 
ElbowPlot(fib_seurat1)
#re-cluster
fib_seurat1 <- FindNeighbors(fib_seurat1, reduction = "pca", dims = 1:50)
fib_seurat1 <- FindClusters(fib_seurat1, verbose = FALSE)
#run umap for visualisation
fib_seurat1 <- RunUMAP(fib_seurat1, reduction = "pca", dims = 1:50)
#this pdf is not used in any figures but shows how granularity within "fib_rich" identity
pdf(paste0(save_path, "fib_seurat1_DimPlot_PC1to50_0.8res.pdf"))
  DimPlot(fib_seurat1, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
#run to identify marker genes within "fib_rich" sub-clusters
fib_seurat1.markers <- FindAllMarkers(fib_seurat1, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(fib_seurat1.markers, paste0(save_path,"fib_seurat1_de_markers_pc50res0.8.csv"), row.names = TRUE)
#assign new names based on gene expression as identified above
DefaultAssay(fib_seurat1)<-"SCT"
fib_seurat1<-SetIdent(fib_seurat1, value = fib_seurat1$seurat_clusters)
new.cluster.ids <- c("fib1","mix","fib3","fib2","fib1","mix","fib4","fib4","fib2","res_mac","fib1","fib3","endothelial","inf_mac1","fib1","fib2","fib4","fib3","inf_mac1","endothelial","mix","fib1","endothelial","mix")
#create new meta data column morpho to retain new names
names(new.cluster.ids) <- levels(fib_seurat1)
fib_seurat1 <- RenameIdents(fib_seurat1, new.cluster.ids)
fib_seurat1$morpho <- Idents(fib_seurat1)
levels(Idents(fib_seurat1))
#this pdf is not used in any figures but shows how granularity within "fib_rich" identity
pdf(paste0(save_path, "fib_seurat1_labelled_DimPlot_PC1to50_0.8res.pdf"))
  DimPlot(fib_seurat1, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
```
SUB_CLUSTER INDIVIDUAL IDENTITIES FROM ABOVE TO FIND GRANULAR PHENOTYPES/CELL IDENTITIES - "fibroblast" 
```{r}
#subset
fib_seurat2 <- subset(x = cosmx, idents = c("fibroblast"))
#run pca
fib_seurat2 <- RunPCA(fib_seurat2, assay = "SCT", verbose = FALSE)
#check pca characteristics 
ElbowPlot(fib_seurat2)
#re-cluster
fib_seurat2 <- FindNeighbors(fib_seurat2, reduction = "pca", dims = 1:50)
fib_seurat2 <- FindClusters(fib_seurat2, verbose = FALSE)
#run umap for visualisation
fib_seurat2 <- RunUMAP(fib_seurat2, reduction = "pca", dims = 1:50)
#this pdf is not used in any figures but shows how granularity within "fibroblast" identity
pdf(paste0(save_path, "fib_seurat2_DimPlot_PC1to50_0.8res.pdf"))
  DimPlot(fib_seurat2, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
#run to identify marker genes within "fibroblast" sub-clusters
fib_seurat2.markers <- FindAllMarkers(fib_seurat2, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(fib_seurat2.markers, paste0(save_path,"fib_seurat2_de_markers_pc50res0.8.csv"), row.names = TRUE)
#assign new names based on gene expression as identified above
DefaultAssay(fib_seurat2)<-"SCT"
fib_seurat2<-SetIdent(fib_seurat2, value = fib_seurat2$seurat_clusters)
new.cluster.ids <- c("fib1","fib2","fib3","fib3","mix","mix","mix","fib1")
#create new meta data column morpho to retain new names
names(new.cluster.ids) <- levels(fib_seurat2)
fib_seurat2 <- RenameIdents(fib_seurat2, new.cluster.ids)
fib_seurat2$morpho <- Idents(fib_seurat2)
levels(Idents(fib_seurat2))
#this pdf is not used in any figures but shows how granularity within "fibroblast" identity
pdf(paste0(save_path, "fib_seurat2_labelled_DimPlot_PC1to50_0.8res.pdf"))
  DimPlot(fib_seurat2, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
```
SUB_CLUSTER INDIVIDUAL IDENTITIES FROM ABOVE TO FIND GRANULAR PHENOTYPES/CELL IDENTITIES - c("macrophage", "neutrophil_rich") 
```{r}
#subset
neutro_seurat <- subset(x = cosmx, idents = c("macrophage", "neutrophil_rich"))
#run pca
neutro_seurat <- RunPCA(neutro_seurat, assay = "SCT", verbose = FALSE)
#check pca characteristics 
ElbowPlot(neutro_seurat)
#re-cluster
neutro_seurat <- FindNeighbors(neutro_seurat, reduction = "pca", dims = 1:50)
neutro_seurat <- FindClusters(neutro_seurat, verbose = FALSE)
#run umap for visualisation
neutro_seurat <- RunUMAP(neutro_seurat, reduction = "pca", dims = 1:50)
#this pdf is not used in any figures but shows how granularity within c("macrophage", "neutrophil_rich") identity
pdf(paste0(save_path, "neutro_seurat_DimPlot_PC1to20_0.8res.pdf"))
  DimPlot(neutro_seurat, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
#run to identify marker genes within c("macrophage", "neutrophil_rich") sub-clusters
neutro_seurat.markers <- FindAllMarkers(neutro_seurat, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(neutro_seurat.markers, paste0(save_path,"neutro_seurat_de_markers_pc50res0.8.csv"), row.names = TRUE)
#assign new names based on gene expression as identified above
DefaultAssay(neutro_seurat)<-"SCT"
neutro_seurat<-SetIdent(neutro_seurat, value = neutro_seurat$seurat_clusters)
new.cluster.ids <- c("MT2A_mac","CHI3L1_mac","CHI3L1_mac","inf_mac2","inf_mac1","CHI3L1_mac","res_mac","inf_mac1","Neutrophil","inf_mac2","inf_mac1","moDC1","CHI3L1_mac","res_mac","mast","mast")
#create new meta data column morpho to retain new names
names(new.cluster.ids) <- levels(neutro_seurat)
neutro_seurat <- RenameIdents(neutro_seurat, new.cluster.ids)
neutro_seurat$morpho <- Idents(neutro_seurat)
levels(Idents(neutro_seurat))
#this pdf is not used in any figures but shows how granularity within c("macrophage", "neutrophil_rich") identity
pdf(paste0(save_path, "neutro_seurat_labelled_DimPlot_PC1to50_0.8res.pdf"))
  DimPlot(neutro_seurat, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
```
USE ALL "LEARNT" UNDER STRUCTURE FROM THE ABOVE SUB-CLUSTERING AND BACK TO THE PARENT OBJECT
PARENT: Cosmx
CHILDREN: "Mac_rich" "mDC" "fib_rich" "fibroblast" c("macrophage", "neutrophil_rich")
```{r}
#get phenotyping from parent - cosmx
meta_cosmx <-cosmx@meta.data
meta_cosmx$final_clustering<-meta_cosmx$nb_clus
meta_cosmx <- meta_cosmx[, c("cell_ID", "final_clustering")]
############################################################
#assign re-clustered IDs from children 
#1 parent, 6 children - "Mac_rich" "mDC" "fib_rich" "fibroblast" c("macrophage", "neutrophil_rich")
############################################################

#get new phenotypying info from child1
meta_mac1 <- mac_seurat1@meta.data
meta_mac1$final_clustering<-meta_mac1$morpho
meta_mac1 <- meta_mac1[, c("cell_ID", "final_clustering")]

#get new phenotyping info from child2
meta_neutro <- neutro_seurat@meta.data
meta_neutro$final_clustering<-meta_neutro$morpho
meta_neutro <- meta_neutro[, c("cell_ID", "final_clustering")]

#get new phenotyping info from child3
meta_mdc <- mDC_seurat@meta.data
meta_mdc$final_clustering<-meta_mdc$morpho
meta_mdc <- meta_mdc[, c("cell_ID", "final_clustering")]

#get new phenotyping info from child4
meta_igg <- igg_seurat@meta.data
meta_igg$final_clustering<-meta_igg$morpho
meta_igg <- meta_igg[, c("cell_ID", "final_clustering")]

#get new phenotyping info from child5
meta_fib1 <- fib_seurat1@meta.data
meta_fib1$final_clustering<-meta_fib1$morpho
meta_fib1 <- meta_fib1[, c("cell_ID", "final_clustering")]

#get new phenotyping info from child6
meta_fib2 <- fib_seurat2@meta.data
meta_fib2$final_clustering<-meta_fib2$morpho
meta_fib2 <- meta_fib2[, c("cell_ID", "final_clustering")]

#get differences from parent and children
cell_ids_needed_from_parent <- setdiff(meta_cosmx$cell_ID, c(meta_mac1$cell_ID,
                                                             meta_neutro$cell_ID,
                                                             meta_mdc$cell_ID,
                                                             meta_igg$cell_ID,
                                                             meta_fib1$cell_ID,
                                                             meta_fib2$cell_ID))
                                       #, meta_tnk$cell_ID))
#subset the above cell IDs from parent
meta_cosmx_subset <- subset(meta_cosmx, cell_ID %in% cell_ids_needed_from_parent)
#bind all meta data together
meta_to_add <- rbind(meta_cosmx_subset,meta_mac1, meta_neutro,meta_mdc, meta_igg, meta_fib1,meta_fib2)


#add new metadata to seuratObj
meta_to_add$cell_ID<-NULL
cosmx <- AddMetaData(cosmx, metadata = meta_to_add)

cosmx <- SetIdent(cosmx, value = cosmx@meta.data$final_clustering)
DimPlot(cosmx, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend() 
levels(cosmx)
###########################################################################################################
```
SUBSET MYELOID CELLS 
```{r plotting myeloid pop}
#subset
cosmx_clusterf <- subset(x = cosmx, idents = c("CHI3L1_mac","inf_mac1", "inf_mac2","MT2A_mac","moDC1","res_mac","Neutrophil", "DC1", "DC2","DC3", "platelets", "pDC", "mast"))
levels(cosmx_clusterf)
#run pca
cosmx_clusterf <- RunPCA(cosmx_clusterf, assay = "SCT", verbose = FALSE)
#check pca results
ElbowPlot(cosmx_clusterf)
#re-cluster
cosmx_clusterf <- FindNeighbors(cosmx_clusterf, reduction = "pca", dims = 1:50)
cosmx_clusterf <- FindClusters(cosmx_clusterf, verbose = FALSE, resolution = 1.2)
#run umap for visualisation
cosmx_clusterf <- RunUMAP(cosmx_clusterf, reduction = "pca", dims = 1:50)
#this pdf is not used in any figures but shows how granularity within myeloid cells
pdf(paste0(save_path, "cosmx_clusterf_DimPlot_pc50_res1.2.pdf"))
  DimPlot(cosmx_clusterf, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
#run to identify marker genes within myeloid cell sub-clusters
cosmx_clusterf.markers <- FindAllMarkers(cosmx_clusterf, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(cosmx_clusterf.markers, paste0(save_path,"cosmx_clusterf_de_markers_pc50_res1.2.csv"), row.names = TRUE)
#assign new names based on gene expression as identified above
DefaultAssay(cosmx_clusterf)<-"SCT"
cosmx_clusterf<-SetIdent(cosmx_clusterf, value = cosmx_clusterf$seurat_clusters)
new.cluster.ids <- c("inf_mac2:CHI3L1_mac","mono_mac1","mono_mac2","inf_mac2:CCL18_mac","moDC","pDC","inf_mac1","inf_mac1","inf_mac2:CHI3L1_mac","res_mac1","moDC","cDC2","inf_mac1","res_mac2","neutrophil","res_mac1","res_mac1","inf_mac2:CCL18_mac","moDC","DC3","res_inf_mac1","DC4","mono_mac2","mono_mac2","mast","mono_mac1","moDC","DC1","inf_mac2:CCL18_mac","moDC","mono_mac1","mono_mac2","DC3","fib1","DC3","DC3","DC3","mono_mac1","mono_mac1","DC3","DC3")
#create new meta data column morpho to retain new names
names(new.cluster.ids) <- levels(cosmx_clusterf)
cosmx_clusterf <- RenameIdents(cosmx_clusterf, new.cluster.ids)
cosmx_clusterf$morpho <- Idents(cosmx_clusterf)
levels(Idents(cosmx_clusterf))
############################################################
#assign re-clustered IDs from children and grandchildren
#1 parent, 1 child
############################################################
#get new phenotypying info from child
meta_cosmx_clusterf <- cosmx_clusterf@meta.data
meta_cosmx_clusterf$cluster<-meta_cosmx_clusterf$morpho
meta_cosmx_clusterf <- meta_cosmx_clusterf[, c("cell_ID", "cluster")]
#parent metadata
meta_all <- cosmx@meta.data
meta_all$cluster<-meta_all$final_clus
meta_all <- meta_all[, c("cell_ID", "cluster")]
#get phenotypying info from parent
cell_ids_needed_from_parent_cosmx <- setdiff(meta_all$cell_ID, meta_cosmx_clusterf$cell_ID)
meta_required_cosmx <- subset(meta_all, cell_ID %in% cell_ids_needed_from_parent_cosmx)
#add metadata
meta_to_add_final <- rbind(meta_cosmx_clusterf, meta_required_cosmx)
cosmx <- AddMetaData(cosmx, metadata = meta_to_add_final)
#visualise
cosmx <- SetIdent(cosmx, value = cosmx@meta.data$cluster)
DimPlot(cosmx, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend() 
levels(cosmx)

```
FILTER NON-MYELOID CELLS BY SUBSETTING & THEN RE-CLUSTER FOR FINAL ANNOTATIONS 
```{r re_cluster_for_final_annotations}
#subset/filter myeloid cells
sub_seurat <- subset(x = cosmx, idents = c("inf_mac2:CHI3L1_mac","mono_mac1", "mono_mac2","inf_mac2:CCL18_mac","moDC","pDC","inf_mac1", "res_mac1", "cDC2","DC3", "res_mac2", "neutrophil", "DC3", "res_inf_mac1","DC4","mast","DC1"))
levels(sub_seurat)
#run pca
sub_seurat <- RunPCA(sub_seurat, assay = "SCT", verbose = FALSE)
#run pca characteristics
ElbowPlot(sub_seurat)
#re-cluster
sub_seurat <- FindNeighbors(sub_seurat, reduction = "pca", dims = 1:50)
sub_seurat <- FindClusters(sub_seurat, verbose = FALSE)
#re-visualise
sub_seurat <- RunUMAP(sub_seurat, reduction = "pca", dims = 1:50)
#this pdf is not used in any figures but shows how granularity within myeloid cells
pdf(paste0(save_path, "subcluster_sub_seurat_DimPlo.pdf"))
  DimPlot(sub_seurat, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend()
dev.off()
#run to identify marker genes within myeloid cell sub-clusters
sub_seurat.markers <- FindAllMarkers(sub_seurat, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(sub_seurat.markers, paste0(save_path,"sub_seurat_de_markers_pc50_res0.8.csv"), row.names = TRUE)
#assign new names based on gene expression as identified above
DefaultAssay(sub_seurat)<-"SCT"
sub_seurat<-SetIdent(sub_seurat, value = sub_seurat$seurat_clusters)
new.cluster.ids <- c("inf_mac2_CCL18_mac","inf_mac2_CHI3L1_mac","moDC2","inf_mac2_CCL18_mac","inf_mac1","pDC","inf_mac1","cDC2","moDC1","res_mac","DC3","inf_mac1","neutrophil","DC1","inf_mac2_CCL18_mac","DC1","moDC3","cDC2","DC1","moDC3","inf_mac2_CCL18_mac","mast","moDC3", "moDC3","inf_mac2_CCL18_mac","DC1","moDC1","inf_mac2_CCL18_mac","mix","moDC3","pDC","moDC3")
#create new meta data column morpho to retain new names
names(new.cluster.ids) <- levels(sub_seurat)
sub_seurat <- RenameIdents(sub_seurat, new.cluster.ids)
sub_seurat$morpho <- Idents(sub_seurat)

#check all phenotypes
levels(sub_seurat)
#sub_seurat_without_mix_renamed - RETAINING MYELOID CELLS ONLY 
cosmx_myeloid <- subset(x = sub_seurat, idents = c("mix"), invert=TRUE)
#check all phenotypes
levels(cosmx_myeloid)

###DO NOT UNCOMMENT THE CODE BELOW UNLESS RE-STARTING ANALYSIS
#saveRDS(cosmx_myeloid, file = "cosmx_myeloid_clusterf.Rds")

############################################################
#assign re-clustered IDs from parent and 1 child
#1 parent, 1 child
############################################################
#get new phenotypying info from child
meta_sub_seurat <- sub_seurat@meta.data
meta_sub_seurat$clusterf<-meta_sub_seurat$morpho
meta_sub_seurat <- meta_sub_seurat[, c("cell_ID", "clusterf")]
#parent metadata
meta_all <- cosmx@meta.data
meta_all$clusterf<-meta_all$cluster
meta_all <- meta_all[, c("cell_ID", "clusterf")]
#get phenotypying info from parent
cell_ids_needed_from_parent_cosmx <- setdiff(meta_all$cell_ID, meta_sub_seurat$cell_ID)
meta_required_cosmx <- subset(meta_all, cell_ID %in% cell_ids_needed_from_parent_cosmx)
#add meta data
meta_to_add_final <- rbind(meta_sub_seurat, meta_required_cosmx)
cosmx <- AddMetaData(cosmx, metadata = meta_to_add_final)
#visualise
cosmx <- SetIdent(cosmx, value = cosmx@meta.data$clusterf)
DimPlot(cosmx, reduction = "umap", label = TRUE, pt.size = 0.8, shuffle = TRUE, label.box = TRUE, repel = TRUE) + NoLegend() 
levels(cosmx)
###DO NOT UNCOMMENT THE CODE BELOW UNLESS RE-STARTING ANALYSIS
#complete object with all annotation
# saveRDS(cosmx, file = "cosmx_clusterf.Rds")
# write.csv(meta_to_add_final, paste0(save_path, "sub_cluster_clusterf.csv"))
#find markers for final full object
cosmx <- readRDS(paste0(study_path, "cosmx_clusterf.Rds"))
cosmx.markers <- FindAllMarkers(cosmx, only.pos = TRUE, min.pct = 0.1, logfc.threshold = 0.1)
write.csv(cosmx.markers, paste0(save_path,"cosmx_final_clusterf.csv"), row.names = TRUE)
```

