
```{r message=FALSE}
library(Seurat)
library(dplyr)
library(ggplot2)
library(patchwork)
library(sqldf)
library(dplyr)
library(RColorBrewer)
library(plotly)
library(spatstat)
library(gplots)
library(corrplot)
library(stringr)
library(VennDiagram)
library(grid)
library(reshape2)
library(ggplot2)
library(ggpubr)
#https://bioconductor.org/packages/release/bioc/vignettes/EnhancedVolcano/inst/doc/EnhancedVolcano.html#plot-the-most-basic-volcano-plot
library(EnhancedVolcano)

study_name<-"human_SLCL_Visium"
#PLEASE ADD YOUR WORKING PATH HERE
study_path<-"your_path_here/"

samples <-c("30i","35i","40i", "53i", "47i", "48i")
sample_label<-c("P1","P2","P3", "P4", "P5", "P6")
#Create a sub-folder in your 
#work path where all outputs
#will be stored
save_path<-paste0(study_path,"R/")
save_on<-1

```

```{r read_RDS}
nd2117SLCL_skin_merged <- readRDS(paste0(save_path, "with47_48_integrated_nd2117slcl_dims15_res0.9_prepsct.rds"))
#while the res and dims are not directly
#utilised here these are kept here such 
#name of outputs can contain this information
res=0.9
dims=15
```

```{r differential_expresion INFERRED BY SEURAT CLUSTERS}
# Start the clock!
ptm <- proc.time()
levels(Idents(nd2117SLCL_skin_merged))

#Find Marker genes 
all_markers <- FindAllMarkers(nd2117SLCL_skin_merged, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, assay = "SCT")

write.csv(all_markers, paste0(save_path, "with47_48_markers_dims",dims,"_res",res,".csv"), row.names = TRUE)

all_markers <- read.csv(paste0(save_path, "with47_48_markers_dims",dims,"_res",res,".csv"), header = TRUE)
# Stop the clock
time_taken<- proc.time() - ptm
print(paste0("Time elapsed: ", sprintf((time_taken[3]/60), fmt = '%#.2f')," minutes"))
```


```{r rename clusters based on region location}
# Start the clock!
ptm <- proc.time()
DefaultAssay(nd2117SLCL_skin_merged)<-"SCT"

Idents(nd2117SLCL_skin_merged)<-"orig.ident"
levels(nd2117SLCL_skin_merged)
new.cluster.ids <- sample_label
#create new meta data column sample_label to retain new names
names(new.cluster.ids) <- levels(nd2117SLCL_skin_merged)
new.cluster.ids
# > new.cluster.ids
#  30i  35i  40i  53i  47i  48i 
# "P1" "P2" "P3" "P4" "P5" "P6" 
nd2117SLCL_skin_merged <- RenameIdents(nd2117SLCL_skin_merged, new.cluster.ids)
nd2117SLCL_skin_merged$sample_label <- Idents(nd2117SLCL_skin_merged)

#Fig 1C
pdf(paste0(save_path, "Fig1C_original_identities.pdf"))
  DimPlot(nd2117SLCL_skin_merged, reduction = "umap", pt.size = 1.5, shuffle = TRUE, group.by = "sample_label", label.size = 6)  
dev.off()

#Supplemnetary_Fig 1C
pdf(paste0(save_path, "Supplemnetary_Fig1C_original_identities.pdf"))
  VlnPlot(nd2117SLCL_skin_merged, features = "nFeature_Spatial", group.by = "sample_label") +
    NoLegend() +
    geom_hline(aes(yintercept = 4104), col = "red", linewidth = 0.3) +
    ggtitle(label = "") +
    ylab(label = "Gene count")
dev.off()

#Fig1D/E/F
Idents(nd2117SLCL_skin_merged) <- nd2117SLCL_skin_merged$morpho
ident_colours =c("mix" ="grey", "My1" = "#008000", "B/fib" ="#CD5C5C",'My2'='#ff9a36','TL'='#B95FBB', 'KC1'='#faf4cf',"Fib"='#F0E68C','KC2'='#E6E6FA', 'B'='#F68282','My3'='#25aff5',"Endothelial" ="#ADD8E6")
my_cols1 <- ident_colours[order(as.character(names(ident_colours)))]

#Fig 1D
pdf(paste0(save_path, "Fig1D_PC1to",dims,"res",res,"labelledIndents.pdf"))
  DimPlot(nd2117SLCL_skin_merged, reduction = "umap", label = TRUE, pt.size = 1.5, shuffle = TRUE, cols = my_cols1,label.box = TRUE, repel = TRUE, label.size = 6) + NoLegend() 
dev.off()

#Fig 1E
pdf(paste0(save_path, "Fig1E_heatmap_allclusters_reindented",dims,"_res",res,".pdf"), 20, 15)
  DefaultAssay(nd2117SLCL_skin_merged) <- "SCT"
  all_markers %>%
      group_by(cluster) %>%
      top_n(n = 10, wt = avg_log2FC) -> top10
  DoHeatmap(nd2117SLCL_skin_merged, features = top10$gene, group.colors = my_cols1) + NoLegend()
dev.off()

#Fig 1F and Supplementary Fig 1D
pdf(paste0(save_path, "Fig1F_SpatialDimPlot_labelledindents_PC1to",dims,"res",res,".pdf"))
  x<-SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, pt.size.factor = 1, image.alpha = 0,label = FALSE, label.size = 3, repel = TRUE, cols = my_cols1) 
  y<-SpatialDimPlot(nd2117SLCL_skin_merged, alpha = c(0.9, 0.9), crop = FALSE, pt.size.factor = 0.9, cols = my_cols1)
  for(i in 1:6){
    print(x[[i]])
  }
  for(i in 1:6){
    print(y[[i]])
  }
dev.off() 

#THE FOLOWING GRAPHS ARE THE BASIS OF Fig 1H-I and Supplementary Fig 1 E-K (left panels) 
Idents(nd2117SLCL_skin_merged) <- nd2117SLCL_skin_merged$morpho
int.genes <- c("IDO1","CD274","KRT14", "CCL18", "JCHAIN", "COL1A1", "CCL19", "CHIT1", "SPARCL1")
for (genes in int.genes){
  #feature <- top.features[spatial_feature]
  pdf(paste0(save_path, "Spatial_feature_int_genes_", genes, ".pdf"))
    x<-SpatialFeaturePlot(nd2117SLCL_skin_merged, features = genes, alpha = c(0, 0), crop = FALSE, pt.size.factor = 1)
    y<-SpatialFeaturePlot(nd2117SLCL_skin_merged, features = genes, alpha = c(0, 5), crop = FALSE, pt.size.factor = 1, min.cutoff = 0)
    #, max.cutoff = 10
    z<-FeaturePlot(nd2117SLCL_skin_merged,reduction = "umap",label = TRUE, pt.size = 1.2, features = genes, min.cutoff = 0)
    k<-VlnPlot(nd2117SLCL_skin_merged,features = genes, cols = my_cols1, y.max = 6)
      print(x[[1]]+y[[1]])
      print(x[[2]]+y[[2]])
      print(x[[3]]+y[[3]])
      print(x[[4]]+y[[4]]) 
      print(x[[5]]+y[[5]])
      print(x[[6]]+y[[6]])
      print(z)
      print(k)

  dev.off()
}

#DATA FOR FIG 1H
my1my2my3 <- subset(nd2117SLCL_skin_merged, subset = morpho %in% c("My1","My2","My3"))
fig1i<-VlnPlot(my1my2my3, features = c("IDO1","CD274"), group.by = "morpho")
#use the following dataframes on Graphpad to gene
ido1_across_my1my2my3<-fig1i[[1]][["data"]]
pdl1_across_my1my2my3<-fig1i[[2]][["data"]]

#THE FOLOWING GRAPHS CONTRIBUTE TO  Supplementary Fig 1 e-k (right panels)
## plotting each cluster separately on 40i
cluster_set <-CellsByIdentities(nd2117SLCL_skin_merged)
## KC1 and KC2
pdf(paste0(save_path, "SpatialDimPlot_clusterbyid_KC1_2.pdf"))
SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, cells.highlight = c(cluster_set[["KC1"]],cluster_set[["KC2"]]), pt.size.factor = 1, image.alpha = 0.5,label = FALSE, label.size = 3, repel = TRUE, images = "X40i", cols.highlight=c("#000000", "#DCDCDC"))
dev.off()
## my1
pdf(paste0(save_path, "SpatialDimPlot_clusterbyid_my1.pdf"))
SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, cells.highlight = cluster_set[["My1"]], pt.size.factor = 1, image.alpha = 0.5,label = FALSE, label.size = 3, repel = TRUE, images = "X40i", cols.highlight=c("#000000", "#DCDCDC"))
dev.off()

## B/fib and B
pdf(paste0(save_path, "SpatialDimPlot_clusterbyid_B_fibB.pdf"))
SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, cells.highlight = c(cluster_set[["B"]],cluster_set[["B/fib"]]), pt.size.factor = 1, image.alpha = 0.5,label = FALSE, label.size = 3, repel = TRUE, images = "X40i", cols.highlight=c("#000000", "#DCDCDC"))
dev.off()

pdf(paste0(save_path, "SpatialDimPlot_clusterbyid_B.pdf"))
SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, cells.highlight = c(cluster_set[["B"]]), pt.size.factor = 1, image.alpha = 0.5,label = FALSE, label.size = 3, repel = TRUE, images = "X40i", cols.highlight=c("#000000", "#DCDCDC"))
dev.off()

## B/fib and Fib
pdf(paste0(save_path, "SpatialDimPlot_clusterbyid_fib_fibB.pdf"))
SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, cells.highlight = c(cluster_set[["Fib"]],cluster_set[["B/fib"]]), pt.size.factor = 1, image.alpha = 0.5,label = FALSE, label.size = 3, repel = TRUE, images = "X40i", cols.highlight=c("#000000", "#DCDCDC"))
dev.off()

## TL
pdf(paste0(save_path, "SpatialDimPlot_clusterbyid_tl.pdf"))
SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, cells.highlight = cluster_set[["TL"]], pt.size.factor = 1, image.alpha = 0.5,label = FALSE, label.size = 3, repel = TRUE, images = "X40i", cols.highlight=c("#000000", "#DCDCDC"))
dev.off()

## My3
pdf(paste0(save_path, "SpatialDimPlot_clusterbyid_my2.pdf"))
SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, cells.highlight = cluster_set[["My2"]], pt.size.factor = 1, image.alpha = 0.5,label = FALSE, label.size = 3, repel = TRUE, images = "X40i", cols.highlight=c("#000000", "#DCDCDC"))
dev.off()

## Endothelial
pdf(paste0(save_path, "SpatialDimPlot_clusterbyid_Endothelial.pdf"))
SpatialDimPlot(nd2117SLCL_skin_merged,crop = FALSE, cells.highlight = cluster_set[["Endothelial"]], pt.size.factor = 1, image.alpha = 0.5,label = FALSE, label.size = 3, repel = TRUE, images = "X40i", cols.highlight=c("#000000", "#DCDCDC"))
dev.off()
# Stop the clock
time_taken<- proc.time() - ptm
print(paste0("Time elapsed: ", sprintf((time_taken[3]/60), fmt = '%#.2f')," minutes"))
```

```{r between_clusters}
###Supplementary Fig. 2A
# Start the clock!
ptm <- proc.time()
#DATA FOR Supplementary Fig. 2A
#My2 – My1;  
my2_rich_markers <- FindMarkers(nd2117SLCL_skin_merged, ident.1 = "My2", ident.2 = "My1")
write.csv(my2_rich_markers, paste0(save_path,"my2_rich_markers.csv"))

#My3 – My12; 
my3_rich_markers <- FindMarkers(nd2117SLCL_skin_merged, ident.1 = "My3", ident.2 = c("My1", "My2"))
write.csv(my3_rich_markers, paste0(save_path,"my3_rich_markers.csv"))
# Stop the clock
time_taken<- proc.time() - ptm
print(paste0("Time elapsed: ", sprintf((time_taken[3]/60), fmt = '%#.2f')," minutes"))
```

```{r enhanced_volcano}
#Supplemental Figure 2
lst_gene_lists <- c("my2_rich_markers", "my3_rich_markers")

for(list in lst_gene_lists){
  pdf(paste0(save_path, list, "_volcano.pdf"),8,12)
  df <- read.csv(paste0(save_path, list, ".csv"))
  df$p_val <- NULL
  colnames(df) <- c("gene","log2FoldChange","pct1","pct2","pvalue")
  df$pvalue[df$pvalue==0]<-5e-305
  rownames(df) <- df$gene

  print(EnhancedVolcano(df,
                      lab = rownames(df),
                      x = 'log2FoldChange',
                      xlim = c(-3, 3),
                      y = 'pvalue',
                      title = list,
                      pCutoff = 10e-4,
                      FCcutoff = 0.5,
                      pointSize = 0.5,
                      labSize = 3.0, 
                      subtitle = "",drawConnectors = TRUE,
                      widthConnectors = 0.5, max.overlaps = Inf,legendLabSize = 10, legendIconSize = 3.0)
  )
  dev.off()
}
```


```{r feature scatter ido1 and pdl1}
mean_pdl1<- mean(nd2117SLCL_skin_merged[["SCT"]]["CD274",])
mean_ido1<- mean(nd2117SLCL_skin_merged[["SCT"]]["IDO1",])
#Fig 3A
pdf(paste0(save_path, "47_48_pdl1_ido1_overall_scatter.pdf"), height = 5, width=5)
 FeatureScatter(nd2117SLCL_skin_merged, feature1 = "IDO1", feature2 = "CD274", cols = rep("black",11), jitter=3) + NoLegend() + geom_vline(aes(xintercept = 1.2), col = "red", linewidth = 0.3) + geom_hline(aes(yintercept = 0.5), col = "red", linewidth = 0.3)
dev.off()
```

```{r create annotation column for ido1 pdl1 spots}
#Data preparation for Fig 3B, 3C and Supplemental Figure 4A
###################################################################
DefaultAssay(nd2117SLCL_skin_merged) <- "SCT"
nd2117SLCL_skin_merged$ido1pdl1 <- "rest_of_the_spots"
nd2117SLCL_skin_merged$spot_id<-rownames(nd2117SLCL_skin_merged@meta.data)

ido1_spots<-WhichCells(nd2117SLCL_skin_merged, expression = IDO1 > 1.1 & CD274 < 0.5)
ido1_pdl1_spots<-WhichCells(nd2117SLCL_skin_merged, expression = IDO1 > 1.1 & CD274 > 0.5)
pdl1_spots<-WhichCells(nd2117SLCL_skin_merged, expression = IDO1 < 1.1 & CD274 > 0.5)

nd2117SLCL_skin_merged@meta.data$ido1pdl1[nd2117SLCL_skin_merged@meta.data$spot_id %in% ido1_spots] <- "ido1_spots"
nd2117SLCL_skin_merged@meta.data$ido1pdl1[nd2117SLCL_skin_merged@meta.data$spot_id %in% ido1_pdl1_spots] <- "ido1_pdl1_spots"
nd2117SLCL_skin_merged@meta.data$ido1pdl1[nd2117SLCL_skin_merged@meta.data$spot_id %in% pdl1_spots] <- "pdl1_spots"
```

```{r spatial plots of ido1 and pdl1}
#Fig 3B 
pdf(paste0(save_path, "47_48_ido1_pdl1_spatialplot_ido11dot1pdldot5.pdf"), height = 5, width=5)  
x<- SpatialPlot(nd2117SLCL_skin_merged, group.by= "ido1pdl1", image.alpha = 0.3, crop = F, pt.size.factor = 1, cols = c("ido1_spots"="#6495ED","ido1_pdl1_spots" = "#FF1493","pdl1_spots"= "#FFA500","rest_of_the_spots" = "#E6E6FA"))
 for(i in 1:6){
    print(x[[i]])
  }
dev.off() 
```

```{r}
nd2117SLCL_skin_merged<- SetIdent(object = nd2117SLCL_skin_merged, value = "ido1pdl1")
ido1pdl1markergenes<- FindAllMarkers(nd2117SLCL_skin_merged,assay = "SCT",min.pct = 0.1,logfc.threshold = 0.1)
write.csv(ido1pdl1markergenes, paste0(save_path, "ido1pdl1markergenes.csv"))
sig_ido1pdl1markergenes <- subset(ido1pdl1markergenes, subset = p_val_adj < 0.05)
write.csv(sig_ido1pdl1markergenes, paste0(save_path, "ido1pdl1markergenespval0.05.csv"))

#DATA BASIS FOR Fig3C
pdf(paste0(save_path, "ido1_pdl1_spots_dot_plots_cytokines.pdf"))
  n<-DotPlot(nd2117SLCL_skin_merged, features = c("CCL18","IL1B","IL24","CXCL9","CXCL14","IL32","TNFRSF6B","IFNGR2","CCL19","CXCL12", "CCL8"), group.by = "ido1pdl1", scale = TRUE) + RotatedAxis() + coord_flip() + theme(axis.title.x=element_blank(),axis.title.y=element_blank())
  print(n)
dev.off()

write.csv(n[["data"]], paste0(save_path, "ido1pdl1_lineplot_data_top50de_for_graphpad_new.csv"))
```

```{r plotting cytokines etc in Fig 3}
##and Fig.3 D-N
interesting_genes<- c("CCL18","IL1B","IL24","CXCL9","CXCL14","IL32","TNFRSF6B","IFNGR2","CCL19","LTB","CCR7","CXCL13","PDGFRB","CXCL12")

for (interesting_gene in interesting_genes){
    pdf(paste0(save_path, "Spatial_feature_", interesting_gene ,".pdf"))
    x<-SpatialFeaturePlot(nd2117SLCL_skin_merged, features = interesting_gene, alpha = c(0, 0), crop = FALSE, pt.size.factor = 1)
    y<-SpatialFeaturePlot(nd2117SLCL_skin_merged, features = interesting_gene, alpha = c(5, 5), crop = FALSE, pt.size.factor = 1)
      print(x[[1]]+y[[1]])
      print(x[[2]]+y[[2]])
      print(x[[3]]+y[[3]])
      print(x[[4]]+y[[4]]) 
      print(x[[5]]+y[[5]])
      print(x[[6]]+y[[6]])

  dev.off()
}
SpatialFeaturePlot(nd2117SLCL_skin_merged, features = "TGFB1", alpha = c(5, 5), crop = FALSE, pt.size.factor = 1, min.cutoff = 0.5, images = "X40i")
SpatialFeaturePlot(nd2117SLCL_skin_merged, features = "TNF", alpha = c(5, 5), crop = FALSE, pt.size.factor = 1, min.cutoff = 0.5, images = "X40i")
SpatialFeaturePlot(nd2117SLCL_skin_merged, features = "IFNG", alpha = c(5, 5), crop = FALSE, pt.size.factor = 1, min.cutoff = 0.5, images = "X40i")

# #########################################################################################################################################################
# 
# 
# for (interleukin in interleukins){
#     pdf(paste0(save_path, "Spatial_feature_", interleukin ,".pdf"))
#     x<-SpatialFeaturePlot(nd2117SLCL_skin_merged, features = interleukin, alpha = c(0, 0), crop = FALSE, pt.size.factor = 1)
#     y<-SpatialFeaturePlot(nd2117SLCL_skin_merged, features = interleukin, alpha = c(5, 5), crop = FALSE, pt.size.factor = 1)
#       print(x[[1]]+y[[1]])
#       print(x[[2]]+y[[2]])
#       print(x[[3]]+y[[3]])
#       print(x[[4]]+y[[4]]) 
#   dev.off()
# }
# 
# SpatialFeaturePlot(nd2117SLCL_skin_merged, features = "IL32", alpha = c(5, 5), crop = FALSE, pt.size.factor = 1, min.cutoff = 1, images = "X40i")
# 
# ##############################################################################################################################################################
# 
# 
# for (tnf in inf_tnf_tgf){
#     pdf(paste0(save_path, "Spatial_feature_", tnf ,".pdf"))
#     x<-SpatialFeaturePlot(nd2117SLCL_skin_merged, features = tnf, alpha = c(0, 0), crop = FALSE, pt.size.factor = 1)
#     y<-SpatialFeaturePlot(nd2117SLCL_skin_merged, features = tnf, alpha = c(5, 5), crop = FALSE, pt.size.factor = 1)
#       print(x[[1]]+y[[1]])
#       print(x[[2]]+y[[2]])
#       print(x[[3]]+y[[3]])
#       print(x[[4]]+y[[4]]) 
#   dev.off()
# }
# 
# 
# 

```

```{r cytokines}
##Supplemental Figure 4A
# cytokines

sig_ido1pdl1markergenes <- read.csv(paste0(save_path, "ido1pdl1markergenespval0.05.csv"))


ccl_genes<-grep(pattern = "^CCL", x = sig_ido1pdl1markergenes$gene, value = TRUE)
cx_genes<-grep(pattern = "^CX", x = sig_ido1pdl1markergenes$gene, value = TRUE)
ccr_genes<-grep(pattern = "^CCR", x = sig_ido1pdl1markergenes$gene, value = TRUE)

chemokines <- c(ccl_genes, cx_genes, ccr_genes)
chemokines <- sort(chemokines)
chemokines <- unique(chemokines)

interleukins<-grep(pattern = "^IL", x =  sig_ido1pdl1markergenes$gene, value = TRUE)
interleukins <- sort(interleukins)
interleukins <- unique(interleukins)

interferons <- grep(pattern = "^IFN", x = sig_ido1pdl1markergenes$gene, value = TRUE)
tnfs <-grep(pattern = "^TNF", x = sig_ido1pdl1markergenes$gene, value = TRUE)
tgfs <-grep(pattern = "^TGF", x = sig_ido1pdl1markergenes$gene, value = TRUE)

inf_tnf_tgf <- c(interferons, tnfs, tgfs)
inf_tnf_tgf <- sort(inf_tnf_tgf)
inf_tnf_tgf <- unique(inf_tnf_tgf)

pdf(paste0(save_path, "Supplemental_Figure_4A_ido1_pdl1_spots_dot_plots_DE_chemokines.pdf"))
  x<-DotPlot(nd2117SLCL_skin_merged, features = unique(chemokines), group.by = "ido1pdl1", scale = TRUE, scale.min = 0, col.min = 0, col.max = 1.5) + RotatedAxis() + coord_flip() + theme(axis.title.x=element_blank(), axis.title.y=element_blank())
  Fig1 <- ggarrange(x,ncol = 1,nrow = 1)
  print(Fig1)
dev.off()

pdf(paste0(save_path, "Suppl_Fig4A_nd2117SLCL_skin_merged_IDO1PDL1_dot_plots_interleukins.pdf"))
  x<-DotPlot(nd2117SLCL_skin_merged, features = interleukins, group.by = "ido1pdl1", scale = TRUE, scale.min = 0, col.min = 0, col.max = 1.5) + RotatedAxis() + coord_flip() + theme(axis.title.x=element_blank(),axis.title.y=element_blank()) 
  Fig1 <- ggarrange(x,ncol = 1,nrow = 1)
  print(Fig1)
dev.off()

pdf(paste0(save_path, "Suppl_Fig4A_nd2117SLCL_skin_merged_IDO1PDL1_dot_plots_DE_inf_tnf_tgf.pdf"))
  x<-DotPlot(nd2117SLCL_skin_merged, features = inf_tnf_tgf, group.by = "ido1pdl1", scale = TRUE, scale.min = 0, col.min = 0, col.max = 1.5) + RotatedAxis() + coord_flip() + theme(axis.title.x=element_blank(),axis.title.y=element_blank()) 
  
  Fig1 <- ggarrange(x,ncol = 1,nrow = 1)
  print(Fig1)
dev.off()

###extracted ICs from significant gene list
immune_checks <- c("CTLA4","PDCDLG2","LAG3","CD40","HAVCR2","LILRB2","LILRB4", "CD86", "CD80", "TNFRSF4", "CD276", "CD28" )

pdf(paste0(save_path, "Suppl_Fig4a_nd2117SLCL_skin_merged_dot_plots_ido1pdl1_checkpoints.pdf"))
  #for(cluster in 0:3){
    x<-DotPlot(nd2117SLCL_skin_merged, features = immune_checks, group.by = "ido1pdl1", scale = TRUE, scale.min = 0, col.min = 0, col.max = 1.5) + RotatedAxis() + coord_flip() + theme(axis.title.x=element_blank(),
                axis.title.y=element_blank())
    Fig <- ggarrange(x,ncol = 1,nrow = 1)
    print(annotate_figure(Fig,top = text_grob(paste0("Immune checkpoints") , color = "black", face = "bold", size = 14)))
dev.off()


###spatial plots of 

```



```{r warning=FALSE}
#DATA FOR Fig 7M venn diagram
# Start the clock!
ptm <- proc.time()
######
#function correlation
######
fn_get_corr_mat <- function(obj, gene_name){
  correlation_df_list<-list()


  mat_count<-as.matrix(obj[['SCT']]@data)

  type<-"spearman"

  #get 'gene of interest' expression of all cells
  count_gene<-mat_count[gene_name,]

  correlation_mat<-matrix(nrow = dim(mat_count)[1], ncol = 2)
  #add rownames
  rownames(correlation_mat) <- rownames(mat_count)
  # Create a nested loop
  for(row in 1:nrow(mat_count)) {
    correlation<-stats::cor.test(count_gene, mat_count[row,], method = type)
    correlation_mat[row,1]<-correlation$estimate
    correlation_mat[row,2]<-correlation$p.value
  }
  correlation_mat<-as.data.frame(correlation_mat)
  colnames(correlation_mat)<-c("corr_estimate","pvalue")
  correlation_mat$gene<-rownames(correlation_mat)
  #discard pvalues greater than 0.05 and correlations below 0.1
  correlation_df<-sqldf::sqldf("SELECT gene, corr_estimate, pvalue
                                FROM correlation_mat
                                WHERE pvalue < 0.05
                                AND corr_estimate > 0.1
                                ORDER BY corr_estimate DESC")

  anti_correlation_df<-sqldf::sqldf("SELECT gene, corr_estimate, pvalue
                                FROM correlation_mat
                                WHERE pvalue < 0.05
                                AND corr_estimate < -0.1
                                ORDER BY corr_estimate ASC")
  #write.table(correlation_mat, file = paste0(exp_path, "All_CD8_T_PBMC_Malat1_Vs_allGene_flu_corr.csv"), sep=",")
  correlation_df_list[[paste0("corr_", gene_name)]]<-correlation_df
  correlation_df_list[[paste0("acorr_", gene_name)]]<-anti_correlation_df
  list_corr_mat <- correlation_df_list
  return(list_corr_mat)
}
#DATA FOR Fig 7m venn diagram; 
#code to run the function for PDL1##################################################
gene_name<-"CD274"

list_corr_mat <-fn_get_corr_mat(nd2117SLCL_skin_merged, gene_name)

print(paste0("Time elapsed: ", sprintf((time_taken[3]/60), fmt = '%#.2f')," minutes"))

head(list_corr_mat[[paste0("corr_", gene_name)]],20)
head(list_corr_mat[[paste0("acorr_", gene_name)]],20)

write.table(list_corr_mat[[paste0("corr_", gene_name)]], file = paste0(save_path,"all_corr_",gene_name,".csv"), sep=",", row.names = FALSE)

write.table(list_corr_mat[[paste0("acorr_", gene_name)]], file = paste0(save_path, "all_acorr_",gene_name,".csv"), sep=",", row.names = FALSE)

rm(list_corr_mat)
####################################################################################
#DATA FOR 7 venn diagram; 
#code to run the function for IDO1 #################################################
gene_name<-"IDO1"

list_corr_mat <-fn_get_corr_mat(nd2117SLCL_skin_merged, gene_name)

head(list_corr_mat[[paste0("corr_", gene_name)]],20)
head(list_corr_mat[[paste0("acorr_", gene_name)]],20)

write.table(list_corr_mat[[paste0("corr_", gene_name)]], file = paste0(save_path,"all_corr_",gene_name,".csv"), sep=",", row.names = FALSE)

write.table(list_corr_mat[[paste0("acorr_", gene_name)]], file = paste0(save_path, "all_acorr_",gene_name,".csv"), sep=",", row.names = FALSE)

rm(list_corr_mat)
####################################################################################

